---
title: "R Notebook"
output: html_notebook
---

Import data 

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(janitor)
library(here)
```


I'm going to start by loading in the 2015 data, having a look at it, and tidying it as best as I can 

```{r}
raw_candy_2015 <- read_excel(here("raw_data/candy_ranking_data/boing-boing-candy-2015.xlsx"))

#these data are extremely long, need to pivot longer 
#also a lot of extra pointless questions at the end can go 

#start simple - clean the names 
clean_names_candy_2015 <- clean_names(raw_candy_2015)

names(clean_names_candy_2015)

clean_candy_2015 <- clean_names_candy_2015 %>% 
  pivot_longer(cols = 4:96, 
               names_to = "candy", 
               values_to = "rating")

clean_candy_2015 <- clean_candy_2015 %>%
  select(timestamp, how_old_are_you, are_you_going_actually_going_trick_or_treating_yourself, candy, rating)

#now we have a long tibble - should we adjust timestamp so that it just has year?

clean_candy_2015 <- clean_candy_2015 %>% 
  select(-timestamp) %>%
  add_column(year = 2015)

head(clean_candy_2015)

#now make names clearer 

clean_candy_2015 <- clean_candy_2015 %>% 
  rename(age = how_old_are_you) %>% 
  rename(trick_or_treat = are_you_going_actually_going_trick_or_treating_yourself)

class(clean_candy_2015$age)

#turning age into an integer introduced some NAs by coercion -- might be worth saving it as a separate tibble 

clean_candy_2015<- clean_candy_2015 %>% 
  mutate(age = as.integer(age))



```

Only remaining issue is that of country - can I infer that the 2015 survey was conducted only in the USA? Best not to do that - I will make it NA when I join the data 


Now onto the 2015 data 
```{r}

#load in the data 
raw_candy_2016 <- read_excel(here("raw_data/candy_ranking_data/boing-boing-candy-2016.xlsx"))


#clean names 
clean_names_candy_2016 <- clean_names(raw_candy_2016)

names(clean_names_candy_2016)

#pivot longer so that candy names is one col and ratings one col 
clean_candy_2016 <- clean_names_candy_2016 %>% 
  pivot_longer(cols = 7:106, 
               names_to = "candy", 
               values_to = "rating")

#select only the columns needed for analysis 
names(clean_candy_2016)
clean_candy_2016 <- clean_candy_2016 %>% 
  select(timestamp, are_you_going_actually_going_trick_or_treating_yourself, your_gender, how_old_are_you, which_country_do_you_live_in, candy, rating)

#rename the columns so that they match 2015 and are easier to read 
clean_candy_2016 <- clean_candy_2016 %>% 
  rename(trick_or_treat = are_you_going_actually_going_trick_or_treating_yourself) %>% 
  rename(gender = your_gender) %>% 
  rename(country = which_country_do_you_live_in) %>% 
  rename(age = how_old_are_you)

#add a year column and remove the timestamp column 
clean_candy_2016 <- clean_candy_2016 %>% 
  select(-timestamp) %>% 
  add_column(year = 2016)


#turn the age data into integers 
clean_candy_2016 <- clean_candy_2016 %>% 
  mutate(age = as.integer(age))


```


Now onto the 2017 data 

```{r}
#as above, begin by loading in using read_excel 

raw_candy_2017 <- read_excel(here("raw_data/candy_ranking_data/boing-boing-candy-2017.xlsx"))


#clean the column names 
clean_names_candy_2017 <- clean_names(raw_candy_2017)


#check the names so I know how to pivot 

names(clean_names_candy_2017)

#pivot longer - can use "starts_with()"

clean_candy_2017 <- clean_names_candy_2017 %>% 
  pivot_longer(cols = starts_with("q6"), 
               names_to = "candy", 
               values_to = "rating")

#select only the cols needed for analysis 

clean_candy_2017 <- clean_candy_2017 %>%
  select(q1_going_out, q2_gender, q3_age, q4_country, candy, rating) %>%
  rename(trick_or_treat = q1_going_out, gender = q2_gender, age = q3_age, country = q4_country) %>%
  add_column(year = 2017)

#finally, turn age into an integer 

clean_candy_2017 <- clean_candy_2017 %>% 
  mutate(age = as.integer(age))

#noticed a problem with my data - in the candy column, every name begins with a "q6" - need to use stringr to replace with nothing

clean_candy_2017 <- clean_candy_2017 %>% 
  mutate(candy = str_replace(candy, "q6_", ""))


```


Now I want to join my tibbles together - using a full join? or a row bind??

```{r}
head(clean_candy_2015)
head(clean_candy_2016)
head(clean_candy_2017)

#need to make candy_2015 in the same order as others 


clean_candy_2015 <- clean_candy_2015 %>% 
  add_column(gender = NA, country = NA) %>% 
  select(trick_or_treat, gender, age, country, candy, rating, year)

#now use bind_rows() to bring all the tibbles together 

candy_clean <- bind_rows(clean_candy_2015, clean_candy_2016, clean_candy_2017)


#cool - that gives us a combined tibble! 

head(candy_clean)
tail(candy_clean)

```

Now need to tidy up some of the values in the cols of my tibble 

```{r}
candy_clean <- candy_clean %>% 
  mutate(rating = str_to_lower(rating))


#check the candy names 

candy_clean <- candy_clean %>% 
  mutate(candy = str_replace(candy, "x100_grand_bar", "100_grand_bar"))


#check that there aren't any repeating candy names
candy_clean %>% 
  distinct(candy)

```



Now need to move on to the country column 

```{r}
candy_clean %>% 
  distinct(country)

#need to tidy these strings up! got a lot of different ways to say the USA here 

candy_clean <- candy_clean %>% 
  mutate(country = str_to_lower(country))

full_candy <- candy_clean %>%
  mutate(country = str_replace_all(country, "^us", "usa")) %>% 
  mutate(country = str_replace_all(country, "usa(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "united s(.*)", "usa")) %>% 
  mutate(country = str_replace_all(country, "u.s(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "merica(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "murica(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "^aus(.*)", "australia")) %>%
  mutate(country = str_replace_all(country, "a tropical island south of the equator", NA_character_)) %>%
  mutate(country = str_replace_all(country, "united kindom", "uk")) %>%
  mutate(country = str_replace_all(country, "neverland", NA_character_)) %>%
  mutate(country = str_replace_all(country, "this one", "usa")) %>%
  mutate(country = str_replace_all(country, "units states", "usa")) %>%
  mutate(country = str_replace_all(country, "the best one - usa", "usa")) %>%
  mutate(country = str_replace_all(country, "cascadia", "usa")) %>%
  mutate(country = str_replace_all(country, "the yoo ess of aaayyyyyy", "usa")) %>%
  mutate(country = str_replace_all(country, "somewhere", NA_character_)) %>%
  mutate(country = str_replace_all(country, "god's country", NA_character_)) %>%
  mutate(country = str_replace_all(country, "there isn't one for old men", NA_character_)) %>%
  mutate(country = str_replace_all(country, "eua", "usa")) %>%
  mutate(country = str_replace_all(country, "merca(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "united(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "trump(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "un(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "the usa", "usa")) %>%
  mutate(country = str_replace_all(country, "insanity lately", "usa")) %>%
  mutate(country = str_replace_all(country, "north carolina", "usa")) %>%
  mutate(country = str_replace_all(country, "'usa", "usa")) %>%
  mutate(country = str_replace_all(country, "new(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "california", "usa")) %>%
  mutate(country = str_replace_all(country, "eua", "usa")) %>%
  mutate(country = str_replace_all(country, "^i pretend(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "ahem(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "sub(.*)", "usa")) %>%
  mutate(country = str_replace_all(country, "see above", NA_character_)) %>%
  mutate(country = str_replace_all(country, "denial", NA_character_)) %>%
  mutate(country = str_replace_all(country, "the republic of usa", "usa")) %>%
  mutate(country = str_replace_all(country, "europe", NA_character_)) %>%
  mutate(country = str_replace_all(country, "earth", NA_character_)) %>% 
  mutate(country = str_replace_all(country, "pittsburgh", "usa")) %>%
  mutate(country = str_replace_all(country, "u.k.", "uk")) %>%
  mutate(country = str_replace_all(country, "husa", "usa")) %>%
  mutate(country = str_replace_all(country, "españa", "spain")) %>%
  mutate(country = str_replace_all(country, "not usa", NA_character_)) %>%
  mutate(country = str_replace_all(country, "^([1-9][0-9].*)$", NA_character_)) %>%
  mutate(country = str_replace_all(country, "susa", "usa")) %>%
  mutate(country = str_replace_all(country, "n. ausa", "usa")) %>%
  mutate(country = str_replace_all(country, "atlantis", NA_character_)) %>%
  mutate(country = str_replace_all(country, "narnia", NA_character_)) %>%
  mutate(country = str_replace_all(country, "fear and loathing", NA_character_)) %>%
  mutate(country = str_replace_all(country, "endland", "uk")) %>%
  mutate(country = str_replace_all(country, "murrika", "usa")) %>%
  mutate(country = str_replace_all(country, "alaska", "usa")) %>%
  mutate(country = if_else(country == "soviet canukstan", NA_character_, country)) %>%
  mutate(country = if_else(country == "canada`", "canada", country)) %>%
  mutate(country = if_else(country == "not usa", NA_character_, country)) %>%
  mutate(country = if_else(country == "ud", NA_character_, country)) %>%
  mutate(country = if_else(country == "canae", "canada", country)) %>%
  mutate(country = if_else(country == "can", "canada", country)) %>%
  mutate(country = str_replace_all(country, "^a$", NA_character_)) %>%
  mutate(country = str_replace_all(country, "^1$", NA_character_)) %>%
  mutate(country = if_else(country == "i don't know anymore", NA_character_, country)) %>%
  mutate(country = str_replace_all(country, "^netherlands$", "the netherlands"))
   
    
full_candy %>% 
  distinct(country)
           

 table(full_candy$country)

#it is done!!!




```

Now I need to deal with the outliers 

```{r}

# we need to look at age 
boxplot(full_candy$age)


library(outliers)


class(full_candy$age)
age_zscores <- outliers::scores(unique(full_candy$age))

full_test <- full_candy %>%
  select(age) %>%
  drop_na()

scores()

median(full_test$age)

mean(full_test$age)

# I will set the age limit at 65 because the data is probably skewed - there will be more younger people taking the survey 

table(full_test$age)

full_candy_no_outliers <- full_candy %>%
  filter(age < 65)



age_zscores

is_age_outlier <- age_zscores > 3 | age_zscores < -3


full_candy <- full_candy %>%
  mutate(is_age_outlier = is_age_outlier)

head(full_candy)


#i'm going to remove any data with age as above 100 - because the z scores stuff isn't working... 

full_candy_sans_outliers <- full_candy %>%
  filter(age < 100)


table(full_candy_sans_outliers$age)
```











export the cleaned data 
```{r}
write_csv(full_candy, "~/data_cleaning_project/task_4/clean_data/full_candy.csv")
```








