---
title: "R Notebook"
output: html_notebook
---

# Task 5 Analysis  


```{r echo=FALSE}

library(tidyverse)
library(here)
```


### First, load in the data  

```{r}
rwa_data <- read_csv("clean_data/rwa_clean.csv")

head(rwa_data)

#its read in respondent as a double - not sure if that will be a problem later on.. probs not

```



## 1. What’s the average RWA score for each gender?

```{r}


rwa_data %>%
  group_by(gender) %>%
  summarise(mean(rwa_score))



```


## 2. What’s the average RWA score for left handed people vs. right handed people.
```{r}

rwa_data %>%
  group_by(hand) %>%
  summarise(mean(rwa_score))

```

## 3.What’s the average family size for each type of childhood? 
```{r}

rwa_data %>%
  group_by(urban) %>%
  summarise(mean(familysize))


```


## 4.What’s the average time to take the test for each education level?
```{r}

rwa_data %>%
  group_by(education) %>%
  summarise(mean(testelapse))

```


## 5. Create a plot of results of question 4.

```{r}
test_elapse_by_ed <- rwa_data %>%
  group_by(education) %>%
  drop_na() %>%
  summarise(test_elapse_by_ed = mean(testelapse))

view(test_elapse_by_ed)

ggplot(test_elapse_by_ed, aes(x = education, y = test_elapse_by_ed)) + geom_bar(stat = "identity")
```



## 6. What’s the average RWA score for people aged

### a. Under 18
```{r}
rwa_data %>%
  filter(age < 18) %>%
  summarise(mean(rwa_score))

rwa_data %>%
  filter(age < 18) %>%
  summarise(n_distinct(respondent))
```

### b. 18-25
```{r}
rwa_data %>%
  filter(between(age, 18, 25)) %>%
  summarise(mean(rwa_score))

rwa_data %>%
  filter(between(age, 18, 25)) %>%
  summarise(n_distinct(respondent))
```


### c. 26 - 40

```{r}
rwa_data %>%
  filter(between(age, 26, 40)) %>%
  summarise(mean(rwa_score))

rwa_data %>%
  filter(between(age, 26, 40)) %>%
  summarise(n_distinct(respondent))
```
### d. 41 - 60
```{r}
rwa_data %>%
  filter(between(age, 41, 60)) %>%
  summarise(mean(rwa_score))

rwa_data %>%
  filter(between(age, 41, 60)) %>%
  summarise(n_distinct(respondent))
```

### e. 60 plus
```{r}
rwa_data %>%
  filter(age > 60) %>%
  summarise(mean(rwa_score))


#how many in this group??

rwa_data %>%
  filter(age > 60) %>%
  summarise(n_distinct(respondent))
```


It seems a bit strange that the authoritarianism score goes down with age... will have a look at other aspects of the data to try to explain why - it may be because the bulk of the respondents are under 30. Median age is 27 and mean is 32.5


could create an age_groups colum 
```{r}

rwa_data %>%
  mutate(age_groups = )

mean(rwa_data$age)
median(rwa_data$age)

table(rwa_data$age)
```


