---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

load in the clean data 
```{r}

decathlon <- read_csv("~/data_cleaning_project/task_1/clean_data/clean_decathlon.csv")
```

1.Who had the longest long jump seen in the data?
```{r}
decathlon %>% 
  filter(events == "long_jump") %>% 
  arrange(desc(scores)) %>%
  top_n(1)


# clay had the longest jump - 7.96 (metres I am assuming)
```



2. What was the average 100m time in each competition?
```{r}
decathlon %>% 
  filter(events == "x100m") %>% 
  group_by(competition) %>% 
  summarise(mean(scores))


```

3. Who had the highest total points across both competitions?
```{r}
decathlon %>% 
  select(athlete_name, points) %>% 
  arrange(desc(points)) %>% 
  top_n(1)
```


4. What was the shot-put scores for the top three competitors in each competition?
```{r}

decathlon %>%
  filter(events == "shot_put") %>%
  group_by(competition) %>%
  arrange(desc(scores)) %>% 
  top_n(3)
  

```


5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
#split into two - so for first group (those with 400m < 50s)
decathlon %>% 
  filter(events == "x400m") %>% 
  filter(scores < 50) %>% 
  summarise(mean(points))


#second group (400m > 50s)

decathlon %>% 
  filter(events == "x400m") %>% 
  filter(scores > 50) %>% 
  summarise(mean(points))
```


















