---
title: "R Notebook"
output: html_notebook
---

Data Cleaning - RWA data 

```{r}
library(tidyverse)
library(janitor)
library(here)

```


Load in the data 
```{r}
rwa_raw <- read_csv(here("raw_data/rwa.csv"))


```

Check dimensions, names etc
```{r}
dim(rwa_raw)

#9881 by 90 - thats a lot of columns - should I try to make the data long??

names(rwa_raw)

rwa_clean_names <- clean_names(rwa_raw)
  
names(rwa_clean_names)
```
Need to add a column with the respondent number before we pivot longer 
```{r}
rwa_numbered <- rwa_clean_names %>%
  rownames_to_column(var = "respondent")
```







Need to rework the reverse order questions a bit before we pivot 

```{r}

names(rwa_numbered)


#write a function that subtracts 10



subtract10 <- function(x){
  10 - x
}

#now apply this function across the columns that need to be reversed - using mutate_at()
rwa_reversed <- rwa_numbered %>%
  mutate_at(.funs = subtract10, .vars = c("q4", "q6", "q8", "q9", "q11", "q13", "q15", "q20", "q21"))

  
```





Now to turn data into long format - pivoting on the authoritarianism questions 
```{r}

rwa_pivot <- rwa_reversed %>%
  pivot_longer(cols = starts_with("q"),
               names_to = "auth_q", 
               values_to = "response")

names(rwa_pivot)

#I only need: respondent, ip_country, education, urban, gender, engnat, age, testelapse, hand, age, religion, voted, race, marries, auth_q, response

rwa_clean <- rwa_pivot %>%
  select(-matches("e[1-9]*[0-9]+")) %>%
  select(-starts_with("vc")) %>%
  select(-starts_with("ti")) %>%
  select(-introelapse, -surveyelapse) %>%
  select(-starts_with("scr")) %>%
  select(-surveyaccurate)


head(rwa_clean)
```

Now want to make the other columns more interpretable 

```{r}
#start with urban -- change to childhood_area 




rwa_clean <- rwa_clean %>%
  mutate(urban = recode(urban, `0` = NA_character_, `1` = "rural", `2` = "suburban", `3` = "urban"))


#now the same with gender 

rwa_clean <- rwa_clean %>%
  mutate(gender = recode(gender, `0` = NA_character_, `1` = "male", `2` = "female", `3` = "other"))

#now hand 
rwa_clean <- rwa_clean %>%
  mutate(hand = recode(hand, `0` = NA_character_, `1` = "right", `2` = "left", `3` = "both"))

#now orientation

rwa_clean <- rwa_clean %>%
  mutate(orientation = recode(orientation, `0` = NA_character_, `1` = "heterosexual", `2` = "bisexual", `3` = "homosexual", `4` = "asexual", `5` = "other"))


#now race 

rwa_clean <- rwa_clean %>%
  mutate(race = recode(race, `0` = NA_character_, `1` = "asian", `2` = "arab", `3` = "black", `4` = "white/indigenous aus & american", `5` = "other"))

#voted in past year

rwa_clean <- rwa_clean %>%
  mutate(voted = recode(voted, `0` = NA_character_, `1` = "yes", `2` = "no"))

#married
rwa_clean <- rwa_clean %>%
  mutate(married = recode(married, `0` = NA_character_, `1` = "never married", `2` = "currently married", `3` = "previously married"))


#religion 

rwa_clean <- rwa_clean %>%
  mutate(religion = recode(religion, `0` = NA_character_, `1` = "agnostic", `2` = "atheist", `3` = "buddhist", `4` = "catholic", `5` = "mormon", `6` = "protestant", `7` = "other christian", `8` = "hindu", `9` = "jewish", `10` = "muslim", `11` = "sikh", `12` = "other"))

#education

rwa_clean <- rwa_clean %>%
  mutate(education = recode(education, `0` = NA_character_, `1` = "less than high school", `2` = "high school", `3` = "university degree", `4` = "graduate degree"))



head(rwa_clean)
```


```{r}
head(rwa_clean)
```

Finally, I need to create a rwa score for each respondent 

```{r}

rwa_clean <- rwa_clean %>%
  group_by(respondent) %>%
  filter(auth_q %in% c("q3", "q4", "q5", "q6", "q7", "q8", "q9", "q10", "q11", "q12", "q13", "q14", "q15", "q16", "q17", "q18", "q19", "q20", "q21", "q22")) %>%
  mutate(rwa_score = mean(response))




view(rwa_clean)

```





I think that is as clean as I need for some analysis 

write to a csv

```{r}
write_csv(rwa_clean, "~/data_cleaning_project/task_5/clean_data/rwa_clean.csv")
```

