---
title: "R Notebook"
output: html_notebook
---
Task 4 analysis 

```{r}
library(tidyverse)


#found i needed to specify col_types when loading in the csv...

full_candy <- read_csv("clean_data/full_candy.csv", col_types = "ccicccd"

head(full_candy)
```






1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}

#need to create a pipe that filters for ratings that have a value (are not na)

#after filtering, try to summarise using n() - which will count the number of rows that have a value in rating - giving us the number of ratings in the data 

full_candy %>%
  filter(!is.na(rating)) %>%
  summarise(n())
```



2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?
```{r}

#group_by needed here for trick or treating 

boxplot(full_candy$age)

mean(full_candy$age, na.rm = TRUE)

#have a horrendous outlier in the ages column!!


#going to create a dataset without outliers 

library(outliers)


class(full_candy$age)
age_zscores <- outliers::scores(unique(full_candy$age))

age_zscores

is_age_outlier <- age_zscores > 3 | age_zscores < -3

full_candy <- full_candy %>%
  mutate(is_age_outlier = is_age_outlier)

head(full_candy)

full_candy %>%
  group_by(trick_or_treat) %>%
  summarise(mean(age, na.rm = TRUE))


full_candy %>% 
  arrange(desc(age))

## go back to question and filter for non-outliers 
full_candy_sans_outliers %>%
  group_by(trick_or_treat) %>%
  summarise(mean(age, na.rm = TRUE))


  

```



3. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
#for Joy
full_candy %>%
  filter(rating == "joy") %>%
  group_by(candy) %>%
  summarise(no_of_joy = n()) %>%
  arrange(desc(no_of_joy)) %>%
  top_n(3)


#for despair 
full_candy %>%
  filter(rating == "despair") %>%
  group_by(candy) %>%
  summarise(no_of_joy = n()) %>%
  arrange(desc(no_of_joy)) %>%
  top_n(3)

#for meh

full_candy %>%
  filter(rating == "despair") %>%
  group_by(candy) %>%
  summarise(no_of_joy = n()) %>%
  arrange(desc(no_of_joy)) %>%
  top_n(3)

```

4. How many people rated starburst as despair? 

```{r}
full_candy %>%
  filter(candy == "starburst") %>%
  filter(rating =="despair") %>%
  summarise(n())
```


Now need to count despair as -1, joy as +1 and meh as 0 
Wrangle 
```{r}

#create a new tibble with an ordinal rating measure 

full_candy_rating_ordinal <- full_candy %>%
  mutate(rating_ordinal = recode(rating, "despair" = -1, "meh" = 0, "joy" = 1))


class(full_candy_rating_ordinal$rating_ordinal)

#as it is numeric - it should be easier to work with 



```


6. What was the most popular candy bar by this rating system for each gender in the dataset?
```{r}

table(full_candy_rating_ordinal$gender)

#I think easiest way is to do this for each gender - filtering for gender each time 

#firstly, for "female"

full_candy_rating_ordinal %>%
  filter(gender == "Female") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)


#then for "i'd rather not say"
full_candy_rating_ordinal %>%
  filter(gender == "I'd rather not say") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)


#for male
full_candy_rating_ordinal %>%
  filter(gender == "Male") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)



```
7. What was the most popular candy bar in each year?

```{r}
#filter for the year, then do as above 

full_candy_rating_ordinal %>%
  filter(year == 2015) %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)


  
```

8.What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
```{r}

#for the US 
full_candy_rating_ordinal %>%
  filter(country == "usa") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)

#for the uk

full_candy_rating_ordinal %>%
  filter(country == "uk") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)

#all other country

full_candy_rating_ordinal %>%
  filter(country != "usa") %>%
  group_by(candy) %>%
  summarise(candy_popularity = sum(rating_ordinal, na.rm = TRUE)) %>%
  arrange(desc(candy_popularity)) %>%
  top_n(1)



```

