---
title: "R Notebook"
output: html_notebook
---
Data cleaning task 3


```{r}
library(tidyverse)
library(readr)

library(readxl)
```

```{r}
ship_data <- read_excel("raw_data/seabirds.xls", sheet = 1)
```

```{r}
here::here()
bird_data <- read_excel("raw_data/seabirds.xls", sheet = 2)

head(ship_data)
head(bird_data)
```

```{r}
seabird_observation <- left_join(ship_data, bird_data, "RECORD ID")
```


```{r}
library(janitor)

seabird_observation <- clean_names(seabird_observation)

view(seabird_observation)
seabird_obs_clean <- seabird_observation %>% 
  select(record_id, lat, long, species_common_name_taxon_age_sex_plumage_phase, count)


```




need to rename that horrible names column 
```{r}
seabird_clean <- seabird_obs_clean %>% 
  rename(species = species_common_name_taxon_age_sex_plumage_phase)
```


and change the cases 
```{r}
seabird_clean <- seabird_clean %>%
  mutate(species = str_to_lower(species))
```

Realised during my analysis that I have a lot of nas -- wondering what to do about them
Going to try to isolate the observations that have no bird recorded 
```{r}
seabird_clean %>%
  filter(species == "[no birds recorded]")

#so to drop i will select only the cases that are not that value 

seabird_clean <- seabird_clean %>%
  filter(species != "[no birds recorded]")
```

What about count as NA? 

Should I impute the values or drop them? 


```{r}
seabird_clean %>% 
  filter(is.na(count))


seabird_clean %>% 
  filter(count == 99999)

boxplot(seabird_clean$count)

seabird_clean %>% 
  select(count) %>% 
```
 
 
In the parts of the analysis where this was a problem - I just removed NA values in my code - that seemed like a reasonable solution 






and export... as I think its clean enough
```{r}

library(here)
write_csv(seabird_clean, here("/clean_data/seabird_clean.csv"))
```



